---
title: "Insects analyses and figures"
author: "Nicolas Djeghri"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prepping the analysis
### Call the packages
```{r}
library(lubridate)
library(cmocean)
library(ecotraj)#Main package for trajectory analysis: using version 1.1.0 here
library(ape)
library(vegan)
```


### Call the data
The data originates from the [Sticky pi github page](https://github.com/sticky-pi/sticky-pi-manuscript) authored by Quentin Geissmann. 
```{r}
data <- read.csv("itc_labels.csv")
metadata <- read.csv("metadata.csv")
```

### Shape the data

This code chunk isolate what we need in the data.
```{r}
#Remove non-insect lines
data <- data[data$type_itc=="Insecta",]
#We will work at family level: remove undetermined families:
data <- data[(data$family_itc=="")==F,]

#Rebuild data:
data <- data.frame(trap_id=data$device_series,
                   trap_setDate=ymd_hms(data$start_datetime_series),
                   capture_date=ymd_hms(data$start_datetime),
                   capture_family=data$family_itc)

#Change metadata a tiny bit:
metadata$start_datetime <- ymd_hms(metadata$start_datetime)
metadata$end_datetime <- ymd_hms(metadata$end_datetime)

#add a "trap/time id":
trapTime_id <- paste(data$trap_id,data$trap_setDate)
data <- cbind(trapTime_id,data)

trapTime_id <- paste(metadata$device,metadata$start_datetime)
metadata <- cbind(trapTime_id,metadata)
```

This code chunk goes wide format building a first "sites" x taxa table by binning by hour and trap.
```{r}
#Bin by trap and hours:
counts <- integer(0)
metadataCounts <- integer(0)
families <- unique(data$capture_family)

for (i in unique(data$trapTime_id)){
  databis <- data[data$trapTime_id==i,c("capture_date","capture_family")]
  start <- metadata$start_datetime[metadata$trapTime_id==i]
  end <- metadata$end_datetime[metadata$trapTime_id==i]
  
  nhours <- floor(as.numeric(end-start,units="hours"))-1
  
  countsbis <- matrix(0,nrow=nhours,ncol=length(families))
  colnames(countsbis) <- families
  
  metadataCountsbis <- data.frame(trapTime_id=rep(i,nhours),
                                  time=start+hours(1:nhours),
                                  bait=rep(metadata$vinegar_bait[metadata$trapTime_id==i],nhours))
  
  for (j in 1:nrow(databis)){
    row <- which(((metadataCountsbis$time+hours(1))>=databis$capture_date[j])&(metadataCountsbis$time<databis$capture_date[j]))
    countsbis[row,databis$capture_family[j]] <- countsbis[row,databis$capture_family[j]]+1
  }
  metadataCounts <- rbind(metadataCounts,metadataCountsbis)
  counts <- rbind(counts,countsbis)
}
```

Finally this chunk of code averages out across traps having the same treatment and on three broad periods to get six (three periods x two treatments) day/night cycles (reducing data sparsity). The choice of the three periods is linked to figure 5 in [Geissmann et al. (2022)](https://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.3001689) cutting the period of "low everything" (period A below) of "high Drosophilidae" (period B below) and of "hgh Figitidae" (period C below).

```{r}
#add big periods to group things:
bigPeriods <- rep("A",nrow(metadataCounts))
bigPeriods[metadataCounts$time>=ymd_hms("2020/09/06 00:00:00")] <- "B"
bigPeriods[metadataCounts$time>=ymd_hms("2020/09/25 00:00:00")] <- "C"
metadataCounts <- cbind(metadataCounts,bigPeriods)

AVcounts <- integer(0)
AVmetadata <- integer(0)
for (i in c("Y","N")){
  for (j in c("A","B","C")){
    #by same "real time"
    selec <- (metadataCounts$bait==i)&(metadataCounts$bigPeriods==j)
    newcounts <- apply(counts[selec,],2,tapply,metadataCounts$time[selec],mean)
    
    #by hours of the day
    hours <- (hour(unique(metadataCounts$time[selec]))-8)%%24#note that here we go local time
    newcounts <- apply(newcounts,2,tapply,hours,mean)
    
    
    AVcounts <- rbind(AVcounts,newcounts)
    AVmetadata <- rbind(AVmetadata,data.frame(bait=rep(i,nrow(newcounts)),
                                              bigPeriods=rep(j,nrow(newcounts)),
                                              hours=as.numeric(rownames(newcounts))))
  }
}
```

### Look a bit at the data
Rank-abundance plots:
```{r fig = TRUE, fig.height=5, fig.width=7, fig.align = "center"}
par(mar=c(8,4,2,2))
barplot(sort(apply(AVcounts,2,mean),decreasing=T),las=2,ylab="Abundance",main="On shaped data")

par(mar=c(8,4,2,2))
barplot(sort(table(data$capture_family),decreasing=T),las=2,ylab="Abundance", main="On raw data")
```

Note that we don't have quite the same rank-abundance plots after shaping the data: this is due to averaging. In raw data we simply plot sum the counted individuals, in the shaped data we represent communities (i.e. the composition at a given time) and average them.

# Go trajectory analysis
Preliminary analyses have shown that cycle metrics are quite unclear with that example (data is probably too sparse). What is however interesting is to "focus on fixed-date trajectories" (in this case it is "more fixed-hour trajectories") i.e. the trajectory described by the community at a given point of the day-night cycle over different cycles (here, three cycles).

### Visualize the cyles
Although we won't work too much on them, we can always visualize the cycles. Because we work on three artificial periods and not a true time series, we need to cheat the system a little bit and close the cycles on themselves:
```{r}
#First reshape the abundance table
AVcountsCycles <- integer(0)
AVmetadataCycles <- integer(0)
for (i in c("Y","N")){
  for (j in c("A","B","C")){
    selec <- (AVmetadata$bait==i)&(AVmetadata$bigPeriods==j)
    AVcountsCycles <- rbind(AVcountsCycles,AVcounts[selec,],AVcounts[(AVmetadata$hours==0)&selec,])
    AVmetadataCycles <- rbind(AVmetadataCycles,AVmetadata[selec,],data.frame(bait=i,bigPeriods=j,hours=24))
  }
}

#Then recompute a distance matrix and go ecotraj format:
d <- vegdist(AVcountsCycles)#Note that we work with percentage difference (Bray-Curtis) distances
sites <- paste(AVmetadataCycles$bait,AVmetadataCycles$bigPeriods)
times <- (AVmetadataCycles$hours)
Trajbis <- defineTrajectories(d,sites=sites,times=times)
Cycles <- extractCycles(Trajbis,24)
```

```{r fig = TRUE, fig.height=5, fig.width=5, fig.align = "center"}
colCycles <- c("sienna1","sienna3","sienna4","skyblue1","skyblue3","skyblue4")
names(colCycles) <- unique(sites)
cyclePCoA(Cycles,sites.colors = colCycles,lwd=2,length=0.1)
```

Brown colors indicate the traps with bait (i.e. the experimental traps) whereas the blue colors indicate the controls. Darker indicate latter sampling. Although it is a bit busy, we see first: there are clear community structure linked to the day/night cycle (oof!): the cycles can be followed and are not bunch confusingly in on place. Second: there is some "long-term" change along axis 2: darker cycles  are on top, lighter cycles on the bottom. Third: there seems to be an effect of the experimental treatment: blue cycles are always further on axis 2 as compared to their brown counterparts.  
Let's now look at the fixed-date trajectories:


###Visualize the fixed date trajectories
We start by getting the fixed-date trajectories in ecotraj format. We can't use the dedicated Cyclical Ecological Trajectory Analysis functions in `ecotraj` because again, we do not exactly work with a time series. However, we can tweak our use of classic Ecological Trajectory Analysis functions to make them do what we need!
```{r}
d <- vegdist(AVcounts)
sites <- paste(AVmetadata$bait,AVmetadata$hours)
times <- c(rep(1:3,each=24),rep(1:3,each=24))
fdT <- defineTrajectories(d,sites=sites,times=times)
fdT_centered <- centerTrajectories(fdT)#same but centered
```

### Making figure 10
```{r}
#graphical parameters:
baseCol <- (cmocean("phase")(24))[c(13:24,1:12)]

colfdT_N <- c(baseCol,rep(rgb(1,1,1,0),24))
colfdT_Y <- c(rep(rgb(1,1,1,0),24),baseCol)
colfdT_All <- c(baseCol,baseCol)
pchfdT_All <- c(rep(16,24),rep(1,24))

#first figure Omega0 space---------
pdf("Figure 10.pdf",width=7,height=7)

layout(matrix(c(rep(6,5),1,2,5,6,6,6,5,6,3,4,5),4,4),widths=c(0.3,1,0.25,1),heights=c(0.1,1,1,0.5))

#not centered---------
par(las=1,mar=c(0.5,0,0,0.5),xpd=NA,xaxt="n",col.lab="white")
trajectoryPCoA(fdT,traj.col=colfdT_N,lwd=2,length=0.1,axes=c(1,2))
mtext("Non-centered",line=0.5,font=4,col="grey30")
mtext("Control",line=4,side=2,las=0,font=4,col="grey30")
par(las=1)
legend(legend=expression(paste(Omega[0]," space")),bty="n",x="bottomright",text.font=4)
mtext(side=3,"(a)",cex=0.8,at=-0.9,xpd=NA,line=-1,font=2)

par(xaxt="s",col.lab="black")
title(ylab="PCoA 2 (10%)")

trajectoryPCoA(fdT,traj.col=colfdT_Y,lwd=2,length=0.1,axes=c(1,2))
mtext("Baited",line=4,side=2,las=0,font=4,col="grey30")
par(las=1)
legend(legend=expression(paste(Omega[0]," space")),bty="n",x="bottomright",text.font=4)
mtext(side=3,"(b)",cex=0.8,at=-0.9,xpd=NA,line=-1,font=2)


#centered------
par(xaxt="n",col.lab="white")
trajectoryPCoA(fdT_centered,traj.col=colfdT_N,lwd=2,length=0.1,axes=c(1,2))
mtext("Centered",line=0.5,font=4,col="grey30")
legend(legend=expression(paste(Omega[0]^Cent," space")),bty="n",x="bottomright",text.font=4)
mtext(side=3,"(c)",cex=0.8,at=-0.95,xpd=NA,line=-1,font=2)

par(xaxt="s",col.lab="black")
title(ylab="PCoA 2 (7%)")

trajectoryPCoA(fdT_centered,traj.col=colfdT_Y,lwd=2,length=0.1,axes=c(1,2))
legend(legend=expression(paste(Omega[0]^Cent," space")),bty="n",x="bottomright",text.font=4)
mtext(side=3,"(d)",cex=0.8,at=-0.95,xpd=NA,line=-1,font=2)

#legend-----
par(mar=c(4.5,6,4,6))
barplot(rep(1,24),col=baseCol,space=0,border=NA,yaxt="n",yaxs="i",xaxs="i",ylim=c(0,1))
axis(side=1,at=seq(0,22,2)+0.5,labels=seq(0,22,2))
mtext("Hour of the day",side=1,line=3,cex=0.8)
box()

#END-----
dev.off()
```
### Distance between fixed-date trajectories

We can now compute dissimilarities (as distances) between the fixed-date trajectories and go in a simplified similarity space
```{r}
dfdT <- trajectoryDistances(fdT)
dfdT_centered <- trajectoryDistances(fdT_centered)
```

### Making figure 11
```{r}
DayNight <- rep(c(rep("Night",5),rep("Day",14),rep("Night",5)),2)
Bait <- c(rep("Y",24),rep("N",24))

colDayNight <- c("gold","darkblue")
names(colDayNight) <- c("Day","Night")
colDayNight <- colDayNight[DayNight]


pdf("Figure 11.pdf",width=8,height=4)
layout(matrix(1:2,1,2))
par(mar=c(4,4,2,1))

pcoafdT <- pcoa(dfdT)
plot(pcoafdT$vectors[,1:2],
     xlab=paste0("PCoA 1 (",round(pcoafdT$values$Relative_eig[1]*100)," %)"),
     ylab=paste0("PCoA 2 (",round(pcoafdT$values$Relative_eig[2]*100)," %)"),
     col=colDayNight,pch=pchfdT_All,
     las=1,asp=1,main="",lwd=2,xaxt="n")
axis(side=1,at=c(-0.6,-0.3,0,0.3),labels=c(-0.6,-0.3,0,0.3))
mtext("Non-centered",line=0.5,font=4,col="grey30")
legend(legend=expression(paste(Omega[fdT]," space")),bty="n",x="bottomright",text.font=4)
legend(legend=c("Day hours","Night hours","Control","Baited"),pch=c(22,22,1,16),pt.bg=c("gold","darkblue","black","black"),pt.lwd=c(1,1,2,2),bty="n",x="bottomleft")
mtext(side=3,"(a)",cex=0.8,at=-1.2,xpd=NA,line=-1,font=2)

pcoafdT_centered <- pcoa(dfdT_centered)
plot(pcoafdT_centered$vectors[,1:2],
     xlab=paste0("PCoA 1 (",round(pcoafdT_centered$values$Relative_eig[1]*100)," %)"),
     ylab=paste0("PCoA 2 (",round(pcoafdT_centered$values$Relative_eig[2]*100)," %)"),
     col=colDayNight,pch=pchfdT_All,
     las=1,asp=1,lwd=2,main="")
mtext("Centered",line=0.5,font=4,col="grey30")
legend(legend=expression(paste(Omega[fdT]^Cent," space")),bty="n",x="bottomright",text.font=4)
mtext(side=3,"(b)",cex=0.8,at=-0.9,xpd=NA,line=-1,font=2)

dev.off()
```
### Finally perform the statistical tests
```{r}
#Homogeneity of variance test--------
VarHomfdT <- betadisper(d=dfdT,g=paste(DayNight,Bait))
VarHomfdT_centered <- betadisper(d=dfdT_centered,g=paste(DayNight,Bait))

anova(VarHomfdT)
anova(VarHomfdT_centered)

#PERMANOVAS-------
permanovafdT <- adonis2(dfdT~DayNight*Bait,by="terms")
permanovafdT_centered <- adonis2(dfdT_centered~DayNight*Bait,by="terms")#technically we can't do this one because variance homogeneity is not respected 

permanovafdT
permanovafdT_centered
```
