---
title: "Shorebirds analyses and figures"
author: "Nicolas Djeghri"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data calling and formating

The shorebird data is produced by the [Réserve Naturelle de la baie de Saint-Brieuc](https://www.reservebaiedesaintbrieuc.com/) and was provided by Anthony Sturbois.
```{r}
data <- read.csv("shorebirds_data.csv",sep=";",dec=",")

#go wide format (assuming missing lines are 0 abundances):
time <-integer(0)
counts <- integer(0)

for (i in unique(data$annee)){
  dataSelec <- data[data$annee==i,]
  for (j in unique(dataSelec$mois)){
    dataSelecBis <- dataSelec[dataSelec$mois==j,]
    countsj <- dataSelecBis$nb
    names(countsj) <- dataSelecBis$TAXO_VERNACUL
    
    counts <- rbind(counts,countsj[unique(data$TAXO_VERNACUL)])
    time <- c(time,i+(j-0.5)/12)
  }
}
colnames(counts) <- unique(data$TAXO_VERNACUL)
counts[is.na(counts)] <- 0

#clean the names
colnames(counts) <- c("Barge rousse","Becasseau maubeche","Becasseau variable",
                      "Bernache cravant","Canard chipeau","Canard colvert",
                      "Canard pilet","Canard siffleur","Canard souchet",
                      "Courlis cendre","Huitrier pie","Macreuse noire",
                      "Pluvier argente","Sarcelle d'hiver","Tadorne de Belon",
                      "Tournepierre a collier","Becasseau sanderling","Grand gravelot")
```

# General description of seasonal community dynamics
### Prepping

Start by calling the necessary packages:
```{r}
library(vegan)
library(ecotraj)#Main package for trajectory analysis: using version 1.1.0 here
library(ape)
library(cmocean)
library(colourvalues)
library(Kendall)
```

Put the object in ecotraj format and make some simple analyses:
```{r}
d <- vegdist(counts)#we take the Bray-Curtis distance

Traj <- defineTrajectories(d,times=time,sites=rep("BIRB!!!",length(time)))
Cycles <- extractCycles(Traj,cycleDuration = 1,
                        minEcolStates = 12,
                        startdate = 5.5/12)
#Note that we make the cycle start at date 5.5/12 (this corresponds to June)
#This is because in that site bird populations are most abundant in winter.

#Compute a "year" for each cycle and
#rename the cycles with something more convenient
yearCycles <- floor(tapply(Cycles$metadata$times,Cycles$metadata$cycles,mean))
Cycles$metadata$cycles <- as.character(yearCycles[Cycles$metadata$cycles])

#compute the distance between cycles
dCycles <- trajectoryDistances(Cycles)
#and PCoA to visualize them
pcoa <- pcoa(dCycles)
```

### Making figure 5
```{r}
colcycles <- colour_values(as.numeric(rownames(pcoa$vectors)),"viridis")
names(colcycles) <- rownames(pcoa$vectors)

pdf("Figure 5.pdf",width=8,height=3.5)

layout(matrix(1:2,1,2))

#Cycles-----------
par(las=1,mar=c(4,4,1,1),xaxt="n")
cyclePCoA(Cycles,length=0.1,cycles.colors=colcycles,print.init.points = TRUE)
par(xaxt="s")
axis(side=1,at=c(-0.6,-0.3,0,0.3,0.6),labels=c(-0.6,-0.3,0,0.3,0.6))
text(c("Summer","Winter"),x=c(-0.6,0.5),y=c(0.4,0.42),pos=c(3,3),col="grey40",font=4)
legend(legend=expression(paste(Omega[0]," space")),bty="n",x="bottomright",text.font=4)
mtext(side=3,"(a)",cex=0.8,at=-1.05,xpd=NA,line=-1,font=2)

#distances between cycles------------
plot(pcoa$vectors[,c(1,2)],las=1,col="white",asp=1,pch=16,
     xlab=paste0("PCoA 1 (",round(pcoa$values$Relative_eig[1]*100)," %)"),
     ylab=paste0("PCoA 2 (",round(pcoa$values$Relative_eig[2]*100)," %)"),
)
text(x=pcoa$vectors[,1],y=pcoa$vectors[,2],rownames(pcoa$vectors),col=colcycles,font=2,xpd=NA)
legend(legend=expression(paste(Omega[C]," space")),bty="n",x="bottomright",text.font=4)
mtext(side=3,"(b)",cex=0.8,at=-0.3,xpd=NA,line=-1,font=2)

#END--------
dev.off()
```
# Time series of cycles descriptors
### Computing
```{r}
#Convexity
Convexity <- cycleConvexity(Traj,cycleDuration = 1,
                            minEcolStates = 12,
                            startdate = 5.5/12)
names(Convexity) <- as.character(yearCycles[names(Convexity)])

#Lengths
Lengths <- trajectoryLengths(Cycles)

#LCdBDs
A <- (-0.5*(as.matrix(dCycles)^2))
G <- (diag(ncol(A))-(rep(1,ncol(A))%*%t(rep(1,ncol(A)))/ncol(A)))%*%
      A%*%
     (diag(ncol(A))-(rep(1,ncol(A))%*%t(rep(1,ncol(A)))/ncol(A)))
LCdBDs <- diag(G)/sum(diag(G))
names(LCdBDs) <- rownames(A)
```

### Making figure 6
```{r}
pdf("Figure 6.pdf",width=4,height=6)

layout(matrix(1:5,5,1),heights=c(1,1,1,1,0.2))
par(mar=c(0.5,4,0.5,1))
plot(as.numeric(rownames(pcoa$vectors)),pcoa$vectors[,1],las=1,type="l",ylab="PCoA 1",xlab="",xaxt="n")
points(as.numeric(rownames(pcoa$vectors)),pcoa$vectors[,1],bg=colcycles,pch=21,cex=2,xpd=NA)
mtext(side=3,"(a)",cex=0.8,at=1997,xpd=NA,line=-1,font=2)

plot(y=Convexity,x=as.numeric(names(Convexity)),las=1,type="l",ylab="Cycle convexity",xlab="",xaxt="n")
points(y=Convexity,x=as.numeric(names(Convexity)),bg=colcycles,pch=21,cex=2,xpd=NA)
mtext(side=3,"(b)",cex=0.8,at=1997,xpd=NA,line=-1,font=2)

plot(y=Lengths$Path,x=as.numeric(rownames(Lengths)),las=1,type="l",ylab="Cycle length",xlab="",xaxt="n")
points(y=Lengths$Path,x=as.numeric(rownames(Lengths)),bg=colcycles,pch=21,cex=2,xpd=NA)
mtext(side=3,"(c)",cex=0.8,at=1997,xpd=NA,line=-1,font=2)

plot(y=LCdBDs*100,x=as.numeric(names(LCdBDs)),type="l",las=1,ylab="'LCdBD' (%)",xlab="",xpd=NA)
points(y=LCdBDs*100,x=as.numeric(names(LCdBDs)),bg=colcycles,pch=21,cex=2,xpd=NA)
mtext(side=3,"(d)",cex=0.8,at=1997,xpd=NA,line=-1,font=2)

dev.off()
```

### Tests for the trends

```{r}
MannKendall(pcoa$vectors[,1])
MannKendall(Convexity)
MannKendall(Lengths$Path)
MannKendall(LCdBDs)
```
